---
title: "Nighttime Lights Analysis"
author: "Data Lab"
date: "4/18/2024"
format:
  html:
    toc: true
    code-fold: true
---

Nighttime lights have become a commonly used resource to estimate changes in local economic activity. This section shows changes in nighttime lights in Myanmar.

# Data

We use nighttime lights data from the VIIRS Black Marble dataset. Raw nighttime lights data requires correction due to cloud cover and stray light, such as lunar light. The Black Marble dataset applies advanced algorithms to correct raw nighttime light values and calibrate data so that trends in lights over time can be meaningfully analyzed. From VIIRS Black Marble, we use data from January 2012 through present—where data is available at a 500-meter resolution.

# Methodology

We extract average nighttime lights within different units in Myanmar. We distinguish lights between lights observed in gas flaring locations and lights in other locations. Oil extraction and production involves gas flaring, which produces significant volumes of light. Separately examining lights in gas flaring and other locations allows distinguishing between lights generated due to oil production versus other sources of human activity. We use data on the locations of gas flaring sites from the [Global Gas Flaring Reduction Partnership](https://www.worldbank.org/en/programs/gasflaringreduction); we remove lights within 5km of gas flaring sites.

# Code Setup

Code to download and aggregate nighttime lights data can be found [here](https://github.com/datapartnership/myanmar-economic-monitor/tree/main/notebooks/nighttime-lights/analysis-2024/01_process_data).

This notebook uses the processed data for analysis.

```{r}
#| warning: false
#| message: false

#### Setup
library(tidyverse)
library(sf)
library(leaflet)
library(leaflet.providers)
library(ggpubr)
library(terra)
library(tidyterra)
library(gtools)
library(readxl)
library(janitor)

data_dir <- "~/Dropbox/World Bank/Side Work/Myanmar Economic Monitor/data"

boundaries_dir <- file.path(data_dir, "Boundaries")
ntl_bm_dir     <- file.path(data_dir, "Nighttime Lights BlackMarble")
sez_dir        <- file.path(data_dir, "SEZ")
gadm_dir       <- file.path(data_dir, "GADM")

theme_manual <- theme_classic2() +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        plot.title = element_text(face = "bold"))

```

# Map of nighttime lights

## Interactive map over time

We first show a map of nighttime lights. Most of the country is dark, with lights concentrated within cities.

```{r}
#| warning: false
#| message: false

## Load boundaries
adm0_sf <- read_sf(file.path(boundaries_dir,
                             "RawData",
                             "mmr_polbnda_adm0_250k_mimu.shp"))

## Load/prep raster
prep_r <- function(year_i){
  r <- rast(file.path(ntl_bm_dir, "FinalData", "VNP46A4_rasters",
                      paste0("VNP46A4_t",year_i,".tif")))
  r <- r %>% mask(adm0_sf)
  r[][r[] == 0] <- NA
  r[] <- log(r[] + 1)
  r[] <- log(r[] + 1)
  return(r)
}

r_2012 <- prep_r(2012)
r_2013 <- prep_r(2013)
r_2014 <- prep_r(2014)
r_2015 <- prep_r(2015)
r_2016 <- prep_r(2016)
r_2017 <- prep_r(2017)
r_2018 <- prep_r(2018)
r_2019 <- prep_r(2019)
r_2020 <- prep_r(2020)
r_2021 <- prep_r(2021)
r_2022 <- prep_r(2022)
r_2023 <- prep_r(2023)

## Make map
pal <- colorNumeric(c("yellow", "orange", "red"), unique(c(r_2012[],
                                                           r_2013[],
                                                           r_2014[],
                                                           r_2015[],
                                                           r_2016[],
                                                           r_2017[],
                                                           r_2018[],
                                                           r_2019[],
                                                           r_2020[],
                                                           r_2021[],
                                                           r_2022[],
                                                           r_2023[])),
                    na.color = "transparent")

leaflet() %>%
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  addRasterImage(r_2012, colors = pal, opacity = 1, group = "2012") %>%
  addRasterImage(r_2013, colors = pal, opacity = 1, group = "2013") %>%
  addRasterImage(r_2014, colors = pal, opacity = 1, group = "2014") %>%
  addRasterImage(r_2015, colors = pal, opacity = 1, group = "2015") %>%
  addRasterImage(r_2016, colors = pal, opacity = 1, group = "2016") %>%
  addRasterImage(r_2017, colors = pal, opacity = 1, group = "2017") %>%
  addRasterImage(r_2018, colors = pal, opacity = 1, group = "2018") %>%
  addRasterImage(r_2019, colors = pal, opacity = 1, group = "2019") %>%
  addRasterImage(r_2020, colors = pal, opacity = 1, group = "2020") %>%
  addRasterImage(r_2021, colors = pal, opacity = 1, group = "2021") %>%
  addRasterImage(r_2022, colors = pal, opacity = 1, group = "2022") %>%
  addRasterImage(r_2023, colors = pal, opacity = 1, group = "2023") %>%
  addLayersControl(
    baseGroups = paste0(2012:2023),
    options = layersControlOptions(collapsed=FALSE)
  )
```

## NTL: Static Map

```{r}
#| warning: false
#| message: false

## Load boundaries
adm0_sf <- read_sf(file.path(boundaries_dir,
                             "RawData",
                             "mmr_polbnda_adm0_250k_mimu.shp"))

## Load/prep raster
prep_r <- function(year_i){
  r <- rast(file.path(ntl_bm_dir, "FinalData", "VNP46A4_rasters",
                      paste0("VNP46A4_t",year_i,".tif")))
  r <- r %>% mask(adm0_sf)
  #r[][r[] == 0] <- NA
  r[] <- log(r[] + 1)
  return(r)
}

r_2023 <- prep_r(2023)

ggplot() +
  geom_spatraster(data = r_2023) +
  scale_fill_gradient2(low = "black",
                       mid = "yellow",
                       high = "red",
                       midpoint = 2,
                       na.value = "transparent") +
  labs(title = "Nighttime Lights: 2023") +
  coord_sf() +
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")

```


## NTL: Percent change from 2019

```{r}
#| warning: false
#| message: false

adm3_sf <- read_sf(file.path(gadm_dir,
                             "rawdata",
                             paste0("gadm41_MMR_",3,".json")))

ntl3_df <- readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                             paste0("adm3", "_", "VNP46A4", ".Rds")))

ntl3_df <- ntl3_df %>%
  dplyr::select(GID_3, date, ntl_bm_sum) %>%
  group_by(GID_3) %>%
  dplyr::mutate(ntl_bm_sum_2019 = ntl_bm_sum[date == 2019]) %>%
  ungroup() %>%
  mutate(ntl_bm_sum_pc19 = (ntl_bm_sum - ntl_bm_sum_2019) / ntl_bm_sum_2019 * 100)

ntl3_df$ntl_bm_sum_pc19[ntl3_df$ntl_bm_sum_pc19 > 100] <- 100
ntl3_df$ntl_bm_sum_pc19[ntl3_df$ntl_bm_sum_pc19 < -100] <- -100

adm3_sf <- adm3_sf %>%
  left_join(ntl3_df, by = "GID_3")

adm3_sf %>%
  dplyr::filter(date %in% 2020:2023) %>%
  ggplot() +
  geom_sf(aes(fill = ntl_bm_sum_pc19)) +
  facet_wrap(~date,
             nrow = 1) +
  labs(fill = c("% Change"),
       title = "Percent Change in Nighttime Lights from 2019: ADM3 Level") +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "forestgreen",
                       midpoint = 0,
                       limits = c(-100, 100),
                       breaks = c(-100, -50, 0, 50, 100),
                       labels = c("< -100", "-50", "0", "50", "> 100")) +
  theme_manual +
  theme_void()

adm3_sf %>%
  dplyr::filter(date %in% 2023) %>%
  ggplot() +
  geom_sf(aes(fill = log(ntl_bm_sum+1) )) +
  scale_fill_gradient2(low = "black",
                       mid = "yellow",
                       high = "red",
                       midpoint = 5) +
  labs(fill = "2023 NTL, Logged") +
  theme_manual +
  theme_void()

adm3_sf %>%
  dplyr::filter(date %in% 2023) %>%
  ggplot() +
  geom_sf(aes(fill = ntl_bm_sum )) +
  scale_fill_gradient2(low = "black",
                       mid = "yellow",
                       high = "red",
                       midpoint = 6500) +
  labs(fill = "2023 NTL") +
  theme_manual +
  theme_void()


```

## Per Capita NTL: Percent change from 2019

In this section, we examine the percent change in per capita nighttime lights from 2019. We use population data from the 2016 census to determine population. For each year, we divide total nighttime lights by population to get per capita nighttime lights.

```{r}
#| warning: false
#| message: false

## Census
census_df <- read_xlsx(file.path(data_dir,
                                 "Census",
                                 "BaselineData_Census_Dataset_SR_District_Township_MIMU_16Jun2016_ENG.xlsx"),
                       sheet = 2,
                       skip = 3)

census_df <- census_df %>%
  clean_names()

census_df <- census_df %>%
  dplyr::filter(!is.na(district_pcode)) %>%
  group_by(district_pcode) %>%
  dplyr::summarise(both_sexes_7 = sum(both_sexes_7)) %>%
  ungroup()

## ADM2 File
adm2_sf <- read_sf(file.path(boundaries_dir,
                             "RawData",
                             "mmr_polbnda_adm2_250k_mimu_1.shp"))

adm2_sf <- adm2_sf %>%
  dplyr::rename(district_pcode = DT_PCODE)

## NTL
ntl2_df <- readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                             paste0("admbound2", "_", "VNP46A4", ".Rds")))

ntl2_df <- ntl2_df %>%
  dplyr::select(DT_PCODE, date, ntl_bm_sum) %>%
  group_by(DT_PCODE) %>%
  dplyr::mutate(ntl_bm_sum_2019 = ntl_bm_sum[date == 2019]) %>%
  ungroup() %>%
  dplyr::rename(district_pcode = DT_PCODE)

## Merge
adm2_sf <- adm2_sf %>%
  left_join(ntl2_df, by = "district_pcode") %>%
  left_join(census_df, by = "district_pcode")

## Cleanup
data_sf <- adm2_sf %>%
  dplyr::mutate(ntl_bm_sum_pc = ntl_bm_sum/both_sexes_7,
                ntl_bm_sum_pc_2019 = ntl_bm_sum_2019/both_sexes_7) %>%
  mutate(pc19 = (ntl_bm_sum_pc - ntl_bm_sum_pc_2019) / ntl_bm_sum_pc_2019 * 100)

data_sf$pc19[data_sf$pc19 > 100] <- 100
data_sf$pc19[data_sf$pc19 < -100] <- -100

## Map
data_sf %>%
  dplyr::filter(date %in% 2020:2023) %>%
  ggplot() +
  geom_sf(aes(fill = pc19)) +
  facet_wrap(~date,
             nrow = 1) +
  labs(fill = c("% Change"),
       title = "Percent Change in per Capita Nighttime Lights from 2019: ADM2 Level") +
  scale_fill_gradient2(low = "red",
                       mid = "white",
                       high = "forestgreen",
                       midpoint = 0,
                       limits = c(-100, 100),
                       breaks = c(-100, -50, 0, 50, 100),
                       labels = c("< -100", "-50", "0", "50", "> 100")) +
  theme_manual +
  theme_void()
```

# Spatial Disparities
```{r, fig.height=3}
#| warning: false
#| message: false

ntl1_df <- readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                             paste0("admbound1", "_", "VNP46A4", ".Rds")))

## Yangon
ntl1_df %>%
  group_by(date) %>%
  dplyr::mutate(ntl_bm_sum_yangon = ntl_bm_sum[ST == "Yangon"]) %>%
  ungroup() %>%

  mutate(pc = (ntl_bm_sum - ntl_bm_sum_yangon)/ntl_bm_sum_yangon*100) %>%

  dplyr::filter(date >= 2019) %>%

  ggplot() +
  geom_histogram(aes(x = pc),
                 fill = "dodgerblue3",
                 color = "black",
                 nbin = 10) +
  facet_wrap(~date, nrow = 1) +
  labs(title = "% Difference in Nightime Lights Relative to Yangon",
       x = "% Difference in Nightime Lights Relative to Yangon",
       y = "N Districts") +
  theme_manual

## Yangon
ntl1_df %>%
  group_by(date) %>%
  dplyr::mutate(ntl_bm_sum_npt = ntl_bm_sum[ST == "Nay Pyi Taw"]) %>%
  ungroup() %>%

  mutate(pc = (ntl_bm_sum - ntl_bm_sum_npt)/ntl_bm_sum_npt*100) %>%

  dplyr::filter(date >= 2019) %>%

  ggplot() +
  geom_histogram(aes(x = pc),
                 fill = "dodgerblue3",
                 color = "black",
                 nbin = 10) +
  facet_wrap(~date, nrow = 1) +
  labs(title = "% Difference in Nightime Lights Relative to Nay Pyi Ta",
       x = "% Difference in Nightime Lights Relative to Nay Pyi Ta",
       y = "N Districts") +
  theme_manual
```



# Trends in NTL in Administrative Zones

```{r}
#| warning: false
#| message: false

adm0_df <- readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                             paste0("adm0", "_", "VNP46A4", ".Rds")))

adm1_df <- readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                             paste0("adm1", "_", "VNP46A4", ".Rds")))

adm2_df <- readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                             paste0("adm2", "_", "VNP46A4", ".Rds")))

adm3_df <- readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                             paste0("adm3", "_", "VNP46A4", ".Rds")))

adm0_df %>%
  mutate(ntl_bm_sum = ntl_bm_sum / 1000) %>%
  mutate(ntl_bm_sum_round = round(ntl_bm_sum, 1)) %>%
  ggplot(aes(x = date, y = ntl_bm_sum)) +
  geom_col(fill = "gray70") +
  geom_text(aes(label = ntl_bm_sum_round)) +
  labs(x = NULL,
       y = "NTL Radiance\n(Values in thousands)",
       title = "Sum of nighttime lights within ADM1 units") +
  theme_manual

adm1_df %>%
  mutate(ntl_bm_sum = ntl_bm_sum / 1000) %>%
  mutate(ntl_bm_sum_round = round(ntl_bm_sum, 1)) %>%
  ggplot(aes(x = date, y = ntl_bm_sum)) +
  geom_col(fill = "gray50") +
  #geom_text(aes(label = ntl_bm_sum_round)) +
  labs(x = NULL,
       y = "NTL Radiance\n(Values in thousands)",
       title = "Sum of nighttime lights within ADM1 units") +
  facet_wrap(~NAME_1) +
  theme_manual

adm1_pc_df <- adm1_df %>%
  mutate(ntl_bm_sum = ntl_bm_sum / 1000) %>%
  mutate(ntl_bm_sum_round = round(ntl_bm_sum, 1)) %>%
  dplyr::filter(date %in% c(2021, 2023)) %>%
  pivot_wider(id_cols = NAME_1,
              names_from = date,
              values_from = ntl_bm_sum) %>%
  mutate(pc = (`2023` - `2021`) / `2021` * 100,
         change = (`2023` - `2021`)) %>%
  mutate(pc_lab = paste0(round(pc, 2), "%"),
         change_lab = round(change, 2))

adm1_pc_df %>%
  ggplot(aes(x = pc,
             y = reorder(NAME_1, pc) )) +
  geom_col(fill = "gray70") +
  geom_text(aes(label = pc_lab)) +
  xlim(-110, 110) +
  labs(x = "% change in nighttime lights from 2021 to 2023",
       y = NULL) +
  theme_manual

adm1_pc_df %>%
  ggplot(aes(x = change,
             y = reorder(NAME_1, change) )) +
  geom_col(fill = "gray70") +
  geom_text(aes(label = change_lab)) +
  xlim(-15, 15) +
  labs(x = "Change in nighttime lights from 2021 to 2023",
       y = NULL) +
  theme_manual

adm1_df %>%
  filter(NAME_1 == "Yangon") %>%
  ggplot(aes(x = date, y = ntl_bm_sum)) +
  geom_col() +
  labs(x = NULL,
       y = "NTL Radiance",
       title = "Sum of nighttime lights within Yangon") +
  theme_manual

adm2_df %>%
  filter(NAME_1 == "Yangon") %>%
  ggplot(aes(x = date, y = ntl_bm_sum)) +
  geom_col() +
  facet_wrap(~NAME_2) +
  labs(x = NULL,
       y = "NTL Radiance",
       title = "Sum of nighttime lights within ADM2 units in Yangon") +
  theme_manual

adm3_df %>%
  mutate(NAME_3_clean = paste(NAME_2, NAME_3)) %>%
  filter(NAME_1 == "Yangon") %>%
  ggplot(aes(x = date, y = ntl_bm_sum)) +
  geom_col() +
  facet_wrap(~NAME_3_clean,
             scales = "free_y") +
  labs(x = NULL,
       y = "NTL Radiance",
       title = "Sum of nighttime lights within ADM3 units in Yangon") +
  theme_manual
```

# Trends in NTL in SEZs

## Average across SEZs and non-SEZ locations

```{r}
#| warning: false
#| message: false

sez_df <- readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                            paste0("admsez", "_", "VNP46A4", ".Rds")))

nonsez_df <- readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                               paste0("adm0_rm_sez", "_", "VNP46A4", ".Rds")))

sez_df <- sez_df %>%
  mutate(Name_id = Name %>%
           as.factor() %>%
           as.numeric()) %>%
  mutate(Name = Name %>%
           str_replace_all(" Industrial Zone",
                           "\nIndustrial Zone") %>%
           str_replace_all(" Economic Zone",
                           "\nEconomic Zone") %>%
           str_replace_all(" Industrial Complex",
                           "\nIndustrial Complex") %>%
           str_replace_all(" New Resort",
                           "\nNew Resort"))

sez_df %>%
  group_by(date) %>%
  dplyr::summarise(ntl_bm_sum = mean(ntl_bm_sum)) %>%
  ungroup() %>%
  mutate(ntl_bm_sum = ntl_bm_sum / 1000) %>%
  mutate(ntl_bm_sum_round = round(ntl_bm_sum, 1)) %>%
  ggplot(aes(x = date, y = ntl_bm_sum)) +
  geom_col(fill = "gray70") +
  geom_text(aes(label = ntl_bm_sum_round)) +
  labs(x = NULL,
       y = "NTL Radiance\n(Values in thousands)",
       title = NULL) + # "Average nighttime lights across special economic zones"
   theme_manual

nonsez_df %>%
  mutate(ntl_bm_sum = ntl_bm_sum / 1000) %>%
  mutate(ntl_bm_sum_round = round(ntl_bm_sum, 1)) %>%
  ggplot(aes(x = date, y = ntl_bm_sum)) +
  geom_col(fill = "gray70") +
  geom_text(aes(label = ntl_bm_sum_round)) +
  labs(x = NULL,
       y = "NTL Radiance\n(Values in thousands)",
       title = NULL) + # "Average nighttime lights across Myanmar,\nexcluding special economic zones"
  theme_manual

```

## Individual SEZs

```{r, fig.height=30}
#| warning: false
#| message: false

sez_df %>%
  #dplyr::filter(Name_id %in% 1:30) %>%
  ggplot(aes(x = date, y = ntl_bm_sum)) +
  geom_col() +
  facet_wrap(~Name,
             scales = "free_y",
             ncol = 4) +
  labs(x = NULL,
       y = "NTL Radiance") +
  theme_manual +
  theme(strip.text = element_text(size = 8))


sez_df %>%
  group_by(Name) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  filter(n == 12) %>%
  dplyr::filter(date %in% c(2021, 2023)) %>%
  pivot_wider(id_cols = Name,
              values_from = ntl_bm_sum,
              names_from = date) %>%
  dplyr::mutate(change = `2023` - `2021`) %>%
  ggplot() +
  geom_vline(aes(xintercept = 0), color = "red") +
  geom_histogram(aes(x = change),
                 fill = "gray80",
                 color = "black") +
  theme_manual +
  labs(x = "Change in nighttime lights radiance",
       y = "N SEZs")

```

# Border Locations

## Average across borders
```{r}
#| warning: false
#| message: false

border_df <- bind_rows(
  readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                    paste0("admborder_1km", "_", "VNP46A4", ".Rds"))) %>%
    mutate(buffer = 1),

  readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                    paste0("admborder_2_5km", "_", "VNP46A4", ".Rds"))) %>%
    mutate(buffer = 2.5),

  readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                    paste0("admborder_5km", "_", "VNP46A4", ".Rds"))) %>%
    mutate(buffer = 5),

  readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                    paste0("admborder_10km", "_", "VNP46A4", ".Rds"))) %>%
    mutate(buffer = 10)
) %>%
  mutate(buffer = factor(buffer))

nonborder_df <- bind_rows(
  readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                    paste0("adm0_rm_border_1km", "_", "VNP46A4", ".Rds"))) %>%
    mutate(buffer = 1),

  readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                    paste0("adm0_rm_border_2_5km", "_", "VNP46A4", ".Rds"))) %>%
    mutate(buffer = 2.5),

  readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                    paste0("adm0_rm_border_5km", "_", "VNP46A4", ".Rds"))) %>%
    mutate(buffer = 5),

  readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                    paste0("adm0_rm_border_10km", "_", "VNP46A4", ".Rds"))) %>%
    mutate(buffer = 10)
) %>%
  mutate(buffer = factor(buffer))

border_df <- border_df %>%
  mutate(border_country = case_when(
    border_town == "Kan Paik Ti" ~ "China",
    border_town == "Lwegel" ~ "China",
    border_town == "Mese" ~ "Thailand",
    border_town == "Myawaddy" ~ "Thailand",
    border_town == "Rihkhawdar" ~ "India",
    border_town == "Thantlang" ~ "India",
    border_town == "Tamu" ~ "India",
    border_town == "Myeik" ~ "Port",
    border_town == "Kawthoung" ~ "Thailand",
    border_town == "Sittwe" ~ "Port",
    border_town == "Maungdaw" ~ "Bangladesh",
    border_town == "Muse" ~ "China",
    border_town == "Chinshwehaw" ~ "China",
    border_town == "Kengtung" ~ "China", ## ?? Not exactly on border. Thailand too?
    border_town == "Tachileik" ~ "Thailand"
  ))

border_df %>%
  group_by(date, buffer) %>%
  dplyr::summarise(ntl_bm_sum = sum(ntl_bm_sum)) %>%
  ungroup() %>%

  ggplot(aes(x = date, y = ntl_bm_sum, color = buffer)) +
  geom_line() +
  labs(color = "Buffer (km)") +
  labs(x = NULL,
       y = "NTL Radiance",
       title = "Nighttime lights across border locations") +
  theme_manual

nonborder_df %>%
  ggplot(aes(x = date, y = ntl_bm_sum, color = buffer)) +
  geom_line() +
  labs(color = "Buffer (km)") +
  labs(x = NULL,
       y = "NTL Radiance",
       title = "Nighttime lights across Myanmar, removing border locations") +
  theme_manual
```

## Average across borders by border country

```{r}
#| warning: false
#| message: false

border_df %>%
  group_by(date, buffer, border_country) %>%
  dplyr::summarise(ntl_bm_sum = sum(ntl_bm_sum)) %>%
  ungroup() %>%

  ggplot(aes(x = date, y = ntl_bm_sum, color = buffer)) +
  geom_line() +
  facet_wrap(~border_country, scales = "free_y") +
  labs(color = "Buffer (km)") +
  labs(x = NULL,
       y = "NTL Radiance",
       title = "Nighttime lights across border locations") +
  theme_manual
```

## Individual borders

```{r}
#| warning: false
#| message: false

border_df %>%
  ggplot(aes(x = date, y = ntl_bm_sum, color = buffer)) +
  geom_line() +
  facet_wrap(~border_town, scales = "free_y") +
  labs(color = "Buffer (km)") +
  labs(x = NULL,
       y = "NTL Radiance",
       title = "Average nighttime lights across border locations") +
  theme_manual
```

# Comparing NTL with RWI

Here we compare nighttime lights with the Facebook Relative Wealth Index

```{r}
#| warning: false
#| message: false

# Load data --------------------------------------------------------------------
rwi_df <- readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                            paste0("admrwi", "_", "VNP46A4", ".Rds")))

rwi_df <- rwi_df %>%
  mutate(ntl_bm_sum_ln = log(ntl_bm_sum + 1))

rwi_df$rwi_q4 <- rwi_df$rwi %>%
  quantcut(q = 4, na.rm = TRUE) %>%
  as.numeric() %>%
  factor()

rwi_df$rwi_q5 <- rwi_df$rwi %>%
  quantcut(q = 5, na.rm = TRUE) %>%
  as.numeric() %>%
  factor()

# Scatterplot ------------------------------------------------------------------
rwi_df %>%
  dplyr::filter(date == 2022) %>%
  ggplot(aes(x = ntl_bm_sum_ln,
             y = rwi)) +
  geom_point() +
  #geom_smooth(method='lm', formula= y~x, se = F, color = "darkorange") +
  labs(x = "Nighttime lights, logged",
       y = "Relative Wealth Index",
       title = "Association between nighttime lights and relative wealth index") +
  theme_classic2()

# Trends -----------------------------------------------------------------------
rwi_df %>%
  group_by(date, rwi_q4) %>%
  dplyr::summarise(ntl_bm_sum = sum(ntl_bm_sum)) %>%
  ungroup() %>%
  ggplot(aes(x = date,
             y = ntl_bm_sum)) +
  geom_col() +
  facet_wrap(~rwi_q4,
             scales = "free_y") +
  labs(x = NULL,
       y = "Nighttime lights radiance",
       title = "Nighttime lights by RWI quantile") +
  theme_classic2() +
  theme(strip.background = element_blank())

```

# Correlation with GDP
```{r}
#| warning: false
#| message: false

#### GDP
gdp_df <- read_xlsx(file.path(data_dir, "GDP", "RawData", "Quarterly GDP_2024.xlsx"),
                    skip = 1)

gdp_df <- gdp_df %>%
  clean_names()

gdp_df <- gdp_df %>%
  dplyr::filter(!is.na(months),
                !is.na(quarter)) %>%
  dplyr::select(-c(paste0("x", 19:21))) %>%
  dplyr::select(-c(paste0("x", 23:29))) %>%
  dplyr::select(-x2021)

gdp_long_df <- gdp_df %>%
  pivot_longer(cols = -c(months, quarter, sub_group, economic_activity)) %>%
  mutate(base_year = name %>%
           str_replace_all("x", "") %>%
           str_replace_all("_.*", "") %>%
           as.numeric(),
         end_year = base_year + 1,
         year = case_when(
           quarter %in% c("Q1", "Q2", "Q3") ~ base_year,
           TRUE ~ end_year
         ))

#### NTL
ntl_df <- bind_rows(
  readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                    paste0("adm0", "_", "VNP46A3", ".Rds"))) %>%
    mutate(unit = "Country"),

  readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                    paste0("adm0_rm_border_10km", "_", "VNP46A3", ".Rds"))) %>%
    mutate(unit = "Country, Excluding Borders"),

  readRDS(file.path(ntl_bm_dir, "FinalData", "aggregated",
                    paste0("admsez", "_", "VNP46A3", ".Rds"))) %>%
    mutate(unit = "SEZs")
)

ntl_df <- ntl_df %>%
  dplyr::mutate(base_year = date %>% year,
                month = date %>% month,
                quarter = case_when(
                  month %in% 4:6 ~ "Q1",
                  month %in% 7:9 ~ "Q2",
                  month %in% 10:12 ~ "Q3",
                  month %in% 1:3 ~ "Q4"
                )) %>%
  group_by(unit, base_year, quarter) %>%
  dplyr::summarise(ntl_bm_sum = mean(ntl_bm_sum)) %>%
  ungroup()

## Merge
ntl_gdp_df <- gdp_long_df %>%
  left_join(ntl_df, by = c("base_year", "quarter")) %>%
  mutate(start_month = case_when(
    quarter == "Q1" ~ 4,
    quarter == "Q2" ~ 7,
    quarter == "Q3" ~ 10,
    quarter == "Q4" ~ 1
  )) %>%
  mutate(date = paste0(year, "-", start_month, "-01") %>% ymd()) %>%
  dplyr::mutate(gdp = value %>% as.numeric())

## Country figure
ntl_gdp_df %>%
  filter(unit %in% "Country") %>%
  group_by(date) %>%
  dplyr::summarise(ntl_bm_sum = mean(ntl_bm_sum),
                   gdp = sum(gdp)) %>%
  ungroup() %>%

  dplyr::mutate(gdp = log(gdp),
                ntl_bm_sum = log(ntl_bm_sum)) %>%
  dplyr::filter(!is.na(gdp),
                !is.na(ntl_bm_sum)) %>%

  ggplot(aes(x = ntl_bm_sum,
             y = gdp)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01) +
  labs(title = NULL, # "Correlation between quarterly GDP and NTL"
       x = "Nighttime lights, log",
       y = "GDP, log") +
  theme_manual

## Country figure
ntl_gdp_df %>%
  filter(unit %in% "Country") %>%
  group_by(date) %>%
  dplyr::summarise(ntl_bm_sum = mean(ntl_bm_sum),
                   gdp = sum(gdp)) %>%
  ungroup() %>%
<<<<<<< HEAD

  dplyr::mutate(gdp = log(gdp),
                ntl_bm_sum = log(ntl_bm_sum)) %>%
  dplyr::filter(!is.na(gdp),
                !is.na(ntl_bm_sum)) %>%

=======

>>>>>>> 3e8ee4f4b408019a25df70f5b7521f5e7c3199a3
  ggplot(aes(x = ntl_bm_sum,
             y = gdp)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01) +
  labs(title = NULL, # "Correlation between quarterly GDP and NTL"
       x = "Nighttime lights, log",
       y = "GDP, log") +
  theme_manual

## Country figure
ntl_gdp_df %>%
  filter(unit %in% "SEZs",
         economic_activity %in% "Manufacturing") %>%
  group_by(date) %>%
  dplyr::summarise(ntl_bm_sum = mean(ntl_bm_sum),
                   gdp = sum(gdp)) %>%
  ungroup() %>%
<<<<<<< HEAD

  dplyr::mutate(gdp = log(gdp),
                ntl_bm_sum = log(ntl_bm_sum)) %>%
  dplyr::filter(!is.na(gdp),
                !is.na(ntl_bm_sum)) %>%

=======

>>>>>>> 3e8ee4f4b408019a25df70f5b7521f5e7c3199a3
  ggplot(aes(x = ntl_bm_sum,
             y = gdp)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01) +
  labs(title = NULL, # "Correlation between quarterly Manufacturing GDP and SEZ NTL"
       x = "SEZ Nighttime lights, log",
       y = "Manufacturing GDP, log") +
  theme_manual

## Country figure
ntl_gdp_df %>%
  filter(unit %in% "Country, Excluding Borders") %>%
  group_by(date) %>%
  dplyr::summarise(ntl_bm_sum = mean(ntl_bm_sum),
                   gdp = sum(gdp)) %>%
  ungroup() %>%
<<<<<<< HEAD

  dplyr::mutate(gdp = log(gdp),
                ntl_bm_sum = log(ntl_bm_sum)) %>%
  dplyr::filter(!is.na(gdp),
                !is.na(ntl_bm_sum)) %>%

=======

>>>>>>> 3e8ee4f4b408019a25df70f5b7521f5e7c3199a3
  ggplot(aes(x = ntl_bm_sum,
             y = gdp)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01) +
  labs(title = NULL, # "Correlation between GDP and non-border NTL"
       x = "Non-Border Nighttime lights, log",
       y = "GDP, log") +
  theme_manual
```


```{r}
#| warning: false
#| message: false

## Country figure
ntl_gdp_df %>%
  filter(unit %in% "SEZs",
         economic_activity %in% "Manufacturing") %>%
  group_by(date) %>%
  dplyr::summarise(ntl_bm_sum = mean(ntl_bm_sum),
                   gdp = sum(gdp)) %>%
  ungroup() %>%

  dplyr::mutate(gdp = log(gdp),
                ntl_bm_sum = log(ntl_bm_sum)) %>%
  dplyr::filter(!is.na(gdp),
                !is.na(ntl_bm_sum)) %>%

  arrange(date) %>%
  dplyr::mutate(gdp = c(diff(gdp), NA),
                ntl_bm_sum = c(diff(ntl_bm_sum), NA))  %>%

  ggplot(aes(x = ntl_bm_sum,
             y = gdp)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01) +
  labs(title = NULL, # "Correlation between quarterly Manufacturing GDP and SEZ NTL\nFirst Difference"
       x = "SEZ Nighttime lights, log",
       y = "Manufacturing GDP, log") +
  theme_manual
```

```{r}
#| warning: false
#| message: false

## Country figure
ntl_gdp_df %>%
  filter(unit %in% "SEZs",
         economic_activity %in% "Manufacturing",
         date < ymd("2020-01-01")) %>%
  group_by(date) %>%
  dplyr::summarise(ntl_bm_sum = mean(ntl_bm_sum),
                   gdp = sum(gdp)) %>%
  ungroup() %>%

  dplyr::mutate(gdp = log(gdp),
                ntl_bm_sum = log(ntl_bm_sum)) %>%
  dplyr::filter(!is.na(gdp),
                !is.na(ntl_bm_sum)) %>%

  arrange(date) %>%
  dplyr::mutate(gdp = c(diff(gdp), NA),
                ntl_bm_sum = c(diff(ntl_bm_sum), NA))  %>%

  ggplot(aes(x = ntl_bm_sum,
             y = gdp)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = F) +
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01) +
  labs(title = NULL, # "Correlation between quarterly Manufacturing GDP and SEZ NTL\nFirst Difference (Using data up to 2020)"
       x = "SEZ Nighttime lights, log",
       y = "Manufacturing GDP, log") +
  theme_manual
```

# Limitations

Nighttime lights are a common data source for measuring local economic activity. However, it is a proxy that is strongly—although imperfectly—correlated with measures of interest, such as population, local GDP, and wealth. Consequently, care must be taken in interpreting reasons for changes in nighttime lights.
